---
title: "Week 2 – Wrangling & Descriptive Stats"
author: "Venu Gopal"
format: html
editor: visual
---

```{r}
# Load dataset from data folder
data <- read.csv("../data/student_registrations.csv")



# Preview first 10 rows
head(data, 10)



```

```{r}
# Inspect dataset structure

str(data)

# Dimensions
nrow(data)   # number of rows
ncol(data)   # number of columns

# Column names
names(data)

```

\

```{r}
## Basic descriptive statistics
# Quick numeric summaries
summary(data)

# Specific checks
mean(data$grade, na.rm = TRUE)
median(data$study_hours, na.rm = TRUE)

# Frequency tables for categorical variables
table(data$gender)
table(data$course)

```

```{r}
## Pre-processing checks
# Count missing values in each column
colSums(is.na(data))

```

```{r}
# Grades should be between 0 and 100
data[data$grade < 0 | data$grade > 100, ]

# Attendance rate should be 0–100
data[data$attendance_rate < 0 | data$attendance_rate > 100, ]

# Study hours should be reasonable (0–80)
data[data$study_hours < 0 | data$study_hours > 80, ]

```

```{r}
unique(data$gender)
unique(data$course)

```

```{r}
invalid_dates <- data[
  !grepl("^\\d{4}-\\d{2}-\\d{2}$", data$enrol_date),
  "enrol_date"
]

unique(invalid_dates)

```

```{r}
hist(data$grade, main="Distribution of Grades", xlab="Grade")
hist(data$study_hours, main="Distribution of Study Hours", xlab="Hours")

```

## 

## Data Notes (Week 2)

-   **Missing values:** Age, attendance_rate, assignments_submitted, and satisfaction_level contain missing values.
-   **Anomalies:** Age includes impossible values (e.g. -5, 200). Grades include values below 0 and above 100.
-   **Messy categories:** Gender labels are inconsistent (e.g. Male, M, female, ?).
-   **Dates:** enrol_date includes malformed or empty entries.
-   **Challenge/question:** Should we remove extreme values or cap them?
